---
title: "get the best cluster number"
author: "xyz"
date: "2021/8/1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Deep longitudinal multiomics pro ﬁ ling reveals two biological seasonal patterns in California

Subsequently, we performed Elbow and Silhouette criteria to identify the optimum number of clusters based on C-means clustering for every single ome as well as all omes combined. The sil-houette coefﬁcient estimates the average distance between clusters. 

[silhouette plot in r](https://stackoverflow.com/questions/33999224/silhouette-plot-in-r)

```{r}
library(Mfuzz)
library(vegan)
library(cluster)
library(ggplot2)
library(factoextra)
```

# silhouette width

```{r}
data <-
  read.table(
    "old_version/time_merge.rpkm.tsv",
    head = T,
    sep = "\t",
    stringsAsFactors = F,
    row.names = 1
  )
rt <- as.matrix(data)
eset <- new("ExpressionSet", exprs = rt)
eset.r <- filter.NA(eset, thres = 0.25)
eset.f <- fill.NA(eset.r, mode = "mean")
eset.f <- filter.std(eset.f, min.std = 0)
eset.s <- standardise(eset.f)
eset.m <- t(as.data.frame(eset.s))

m <- mestimate(eset.s)
dis <- dist(eset.m) ^ 2
meanSilWidth <- c()
for (c in 2:20) {
  cl <- mfuzz(eset.s, c = c, m = m)
  sil <- silhouette (cl$cluster, dis)
  meanSilWidth <- c(meanSilWidth, mean(sil[, 3]))
}

df<-data.frame(x=1:20,y=c(0,meanSilWidth))
p<-ggplot(df,aes(x=x,y=y))+
  geom_line()+
  geom_point()+
  xlab("Number of clusters")+
  ylab("Average silhouette width")+
  theme(
    text = element_text(size = 15),
    panel.grid = element_blank(),
    panel.background = element_blank(),
    panel.border = element_rect(
      fill = NA,
      color = "grey10",
      linetype = 1,
      size = 1.
    ))
p+ggsave("Average silhouette width.pdf",width = 8,height = 6)
```

# kmeans elbow silhouette

[Fuzzy cMeans clustering of RNAseq data using mFuzz](https://2-bitbio.com/post/fuzzy-cmeans-clustering-of-rnaseq-data-using-mfuzz/)

[k-means elbow method](https://statsandr.com/blog/clustering-analysis-k-means-and-hierarchical-clustering-by-hand-and-in-r/)

[Deep longitudinal multiomics pro ﬁ ling reveals two biological seasonal patterns in   California](https://github.com/aametwally/ipop_seasonal/blob/master/code/clustering.R)

NbClust

```{r}
Dmin(eset.s, m=m, crange=seq(2,22,1), repeats=3, visu=TRUE)
p<-fviz_nbclust(eset.m, kmeans, method = "wss")
p+ggsave("kmeans elbow.pdf",width = 8,height = 6)
p<-fviz_nbclust(eset.m, kmeans, method = "silhouette")
p+ggsave("kmeans silhouette.pdf",width = 8,height = 6)
```

