---
title: "only reviewed swiss-protein"
author: "xyz"
date: "2021/6/22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# deal with output of maxquant

Identification type: Indicates whether this experiment was identified by MS/MS 
or only by matching between runs.
  
Reverse: When marked with '+', this particular protein group contains noprotein,
made up of at least 50% of the peptides of the leading protein, 
with a peptide derived from the reversed part of the decoy database.
These should be removed for further data analysis. 

```{r}
library(data.table)
library(stringr)
library(VennDiagram)
library(vegan)
library(ggplot2)
library(scales)
library(pheatmap)

# library(extrafont)
# fonts()
# # Show entire table
# fonttable()

df<-read.csv("../only use Swiss-prot proteins/proteinGroupsFiltered.csv", check.names = F)
nrow(df)
# identified 4476 proteins
length(unlist(str_split(df$`Protein IDs`, ";")))

#### Venn diagram ####

COVID <- df$`Identification type COVID1` == "By MS/MS" |
  df$`Identification type COVID2` == "By MS/MS" |
  df$`Identification type COVID3` == "By MS/MS"
CTRL <- df$`Identification type ctrl1` == "By MS/MS" |
  df$`Identification type ctrl2` == "By MS/MS" |
  df$`Identification type ctrl3` == "By MS/MS"
S <- df$`Identification type S1` == "By MS/MS" |
  df$`Identification type S2` == "By MS/MS" |
  df$`Identification type S3` == "By MS/MS"
p <- venn.diagram(
  x = list(
    COVID = (1:nrow(df))[COVID],
    CTRL = (1:nrow(df))[CTRL],
    S = (1:nrow(df))[S]
  ),
  filename = NULL,
  output = F ,
  # imagetype="png" ,
  height = 1200 ,
  width = 1200 ,
  resolution = 30,
  compression = "lzw",
  lwd = 1,
  col = c("#440154ff", '#21908dff', '#fde725ff'),
  fill = c(
    alpha("#440154ff", 0.3),
    alpha('#21908dff', 0.3),
    alpha('#fde725ff', 0.3)
  ),
  cex = 2,
  fontfamily = "sans",
  cat.cex = 2,
  cat.default.pos = "outer",
  cat.pos = c(-27, 27, 135),
  cat.dist = c(0.055, 0.055, 0.085),
  cat.fontfamily = "sans",
  cat.col = c("#440154ff", '#21908dff', '#fde725ff'),
  rotation = 1
)
pdf(file = "identified proteins venn.pdf")
grid.draw(p)
dev.off()

#### heatmap ####

# MS2 count

mat <- df[, 93:101]
mat <- sapply(mat, as.numeric)
mat <- mat[rowSums(mat) != 0, ]
mat <- prop.table(mat,2)
colnames(mat) <-
  c(paste0(rep(c("COVID", "CTRL", "S"), each = 3), c(1:3, 1:3)))
colgroup <-
  data.frame(Treatments = factor(rep(c("COVID", "CTRL", "S"), each = 3)))
row.names(colgroup) <- colnames(mat)

pdf("protein MS2 count heatmap.pdf",
    width = 8,
    height = 6)
pheatmap(
  mat,
  show_rownames = F,
  fontsize = 15,
  # RdBu RdYlBu
  color = colorRampPalette(rev(
    RColorBrewer::brewer.pal(n = 7, name = "RdYlBu")
  ))(100),
  cluster_row = T,
  cluster_col = T,
  scale = "row",
  annotation_col = colgroup
)
dev.off()

#### PCA ####

mat <- df[, 93:101]
mat <- sapply(mat, as.numeric)
mat <- prop.table(mat, margin = 2)
mat <- mat[rowSums(mat) != 0, ]

pca <- rda(t(mat), scale = T)
importance <- summary(pca)[["cont"]][["importance"]]
siteScore <- summary(pca)[["sites"]]
tempDf <- data.frame(x = siteScore[, 1],
                     y = siteScore[, 2],
                     Treatments = colgroup$Treatments)
ggplot(tempDf,
       aes(x = x,
           y = y,
           color = Treatments)) +
  geom_point(size = 3) +
  xlab(paste0("PC1(", round(importance[2, 1] * 100, 2), "%)")) +
  ylab(paste0("PC2(", round(importance[2, 2] * 100, 2), "%)")) +
  theme(text = element_text(size = 20)) +
  ggsave("protein MS2 PCA.pdf",
         width = 8,
         height = 6)
```

# protein with different abundance

```{r}
library(DESeq2)

#### COVID vs CTRL ####
mat <- df[, 93:98]
row.names(mat) <- df$`Majority protein IDs`
mat <- mat[rowSums(mat) != 0, ]
coldata <- data.frame(row.names = colnames(mat),
                      condition = factor(colgroup$Treatments[1:6]))
dds <- DESeqDataSetFromMatrix(countData = mat,
                              colData = coldata,
                              design =  ~ condition)
dds2 <- DESeq(dds,fitType = "local")
DEGs <- results(dds2, contrast = c("condition", "COVID", "CTRL"))
DEGsDf <- na.exclude(as.data.frame(DEGs))
saveRDS(DEGsDf, "COVID vs CTRL.rds")
write.csv(DEGsDf, "COVID vs CTRL.csv")
DEGsDf <-
  readRDS("COVID vs CTRL.rds")
# The Standard of significance was fold change >= 2,
# or =< 1/2 (log(DEGsDlog2FoldChangef)>=1 or <=-1),
# adjusted P value =< 0.05
judgeSignificance <- function(x, y) {
  if (x <= 0.05) {
    if (y <= -1)
      return ("down")
    else if (y >= 1)
      return ("up")
    else
      return ("no")
  }
  else
    return("no")
}
DEGsDf$significance <-
  mapply(judgeSignificance, DEGsDf$padj, DEGsDf$log2FoldChange)
DEGsDfFiltered <- DEGsDf[DEGsDf$significance != "no", ]
write.csv(DEGsDfFiltered,
          "Protein with different abundance COVID vs CTRL.csv")
# minimal padj is 0
head(sort(log10(DEGsDf$padj)))
# change to none zero minimal
DEGsDf$padj[DEGsDf$padj == 0] <- min(DEGsDf$padj[DEGsDf$padj > 0])
drawVolcanoPlot <- function(df) {
  colnames(df) <- c("FDR", "FC")
  df$size <- -log10(df$FDR)
  df <-
    cbind(df, significance = factor(mapply(judgeSignificance, df$FDR, df$FC)))
  ggplot(df, aes(FC,-1 * log10(FDR))) +
    geom_point(
      aes(
        fill = significance,
        colour = significance,
        size = size
      ),
      shape = 21,
      alpha = 0.6
    ) +
    # Resize the point from 0.1 to 10
    scale_size(range = c(.1, 10)) +
    theme(text = element_text(size = 20)) +
    guides(
      fill = guide_legend(override.aes = list(size = 10)),
      size = FALSE,
      colour = FALSE
    ) +
    scale_fill_discrete(labels = paste(levels(df$significance), table(df$significance))) +
    labs(x = expression(log[2](FC)), y = expression(-log[10](FDR))) +
    geom_hline(yintercept = -log10(0.05), linetype = 4) +
    geom_vline(xintercept = c(-log2(2), log2(2)), linetype = 4) +
    # Limit the scope of the drawing, make it left-right symmetric, and add linear coordinates
    scale_x_continuous(limits = c(-31, 31), breaks = sort(c(seq(
      -30,
      30, length.out = 6
    ), -1, 1)))
}

p <- drawVolcanoPlot(DEGsDf[, c("padj", "log2FoldChange")])
# Limit the scope of the drawing and add linear coordinates
p <-
  p + scale_y_continuous(limits = c(0, 280), breaks = c(1.3, 10, 20, 40, 80, 160))
p + ggsave("COVID vs CTRL volcano plot.pdf",
           width = 8,
           height = 6)

#### S vs CTRL ####

mat <- df[, c(99:101, 96:98)]
row.names(mat) <- df$`Majority protein IDs`
mat <- mat[rowSums(mat) != 0, ]
coldata <- data.frame(row.names = colnames(mat),
                      condition = factor(colgroup$Treatments[c(7:9, 4:6)]))
dds <- DESeqDataSetFromMatrix(countData = mat,
                              colData = coldata,
                              design =  ~ condition)
dds2 <- DESeq(dds,fitType = "local")
DEGs <- results(dds2, contrast = c("condition", "S", "CTRL"))
DEGsDf <- na.exclude(as.data.frame(DEGs))
saveRDS(DEGsDf, "S vs CTRL.rds")
write.csv(DEGsDf, "S vs CTRL.csv")
DEGsDf <-
  readRDS("S vs CTRL.rds")
DEGsDf$significance <-
  mapply(judgeSignificance, DEGsDf$padj, DEGsDf$log2FoldChange)
DEGsDfFiltered <- DEGsDf[DEGsDf$significance != "no", ]
write.csv(DEGsDfFiltered, "Protein with different abundance S vs CTRL.csv")
head(sort(log10(DEGsDf$padj)))
DEGsDf$padj[DEGsDf$padj == 0] <- min(DEGsDf$padj[DEGsDf$padj > 0])
p <- drawVolcanoPlot(DEGsDf[, c("padj", "log2FoldChange")])
# Limit the scope of the drawing and add linear coordinates
p <-
  p + scale_y_continuous(limits = c(0, 162),
                         breaks = c(1.3, 10, 20, 40, 80, 160))
# scale_y_break conflict with geom_hline and geom_vline
# p<-p+ scale_y_break(c(40, 150))
p + ggsave("S vs CTRL volcano plot.pdf",
           width = 8,
           height = 6)

#### COVID vs S ####

mat <- df[,c(93:95, 99:101)]
row.names(mat) <- df$`Majority protein IDs`
mat <- mat[rowSums(mat) != 0, ]
coldata <- data.frame(row.names = colnames(mat),
                      condition = factor(colgroup$Treatments[c(1:3,7:9)]))
dds <- DESeqDataSetFromMatrix(countData = mat,
                              colData = coldata,
                              design =  ~ condition)
dds2 <- DESeq(dds,fitType = "local")
DEGs <- results(dds2, contrast = c("condition", "COVID", "S"))
DEGsDf <- na.exclude(as.data.frame(DEGs))
saveRDS(DEGsDf, "COVID vs S.rds")
write.csv(DEGsDf, "COVID vs S.csv")
DEGsDf <-
  readRDS("COVID vs S.rds")
judgeSignificance <- function(x, y) {
  if (x <= 0.05) {
    if (y <= -1)
      return ("down")
    else if (y >= 1)
      return ("up")
    else
      return ("no")
  }
  else
    return("no")
}
DEGsDf$significance <-
  mapply(judgeSignificance, DEGsDf$padj, DEGsDf$log2FoldChange)
DEGsDfFiltered <- DEGsDf[DEGsDf$significance != "no", ]
write.csv(DEGsDfFiltered,
          "Protein with different abundance COVID vs S.csv")
# minimal padj is 0
head(sort(log10(DEGsDf$padj)))
# change to none zero minimal
DEGsDf$padj[DEGsDf$padj == 0] <- min(DEGsDf$padj[DEGsDf$padj > 0])
p <- drawVolcanoPlot(DEGsDf[, c("padj", "log2FoldChange")])
p <-
  p + scale_y_continuous(limits = c(0, 220), breaks = c(1.3, 10, 20, 40, 80, 160))
p + ggsave("COVID vs S volcano plot.pdf",
           width = 8,
           height = 6)
```

# get protein annotation

```{r}
library(AnnotationHub)
library(clusterProfiler)
library(stringr)

# refreshHub(hubClass="AnnotationHub")
ah <- AnnotationHub()
# job::job({display(ah)})
# Homo sapiens OrgDb
hm <- ah[["AH92581"]]
hmDf <-
  AnnotationDbi::select(hm,
                        keys = keys(hm),
                        columns = c("UNIPROT", "GO", "PATH", "ENZYME"))
hmDf <- hmDf[!is.na(hmDf$UNIPROT), ]
goName <-
  AnnotationDbi::select(
    x = GO.db::GO.db,
    keys = hmDf$GO,
    keytype = "GOID",
    columns = "TERM"
  )

#### COVID ####

covid <-
  read.csv("Protein with different abundance COVID vs CTRL.csv")
covidList <- unlist(str_split(covid$X, ";"))
covidList <- unlist(str_split(covidList, "\\|", simplify = T)[, 2])

goMF <-
  enrichGO(
    gene = covidList,
    OrgDb = hm,
    keyType = "UNIPROT",
    ont = "MF",
    pvalueCutoff = 1
  )
goMFdf <- as.data.frame(goMF)
write.csv(goMFdf, "COVID vs CTRL MF enrich.csv", row.names = F)
p <- dotplot(filter(goMF,qvalue<0.05), font.size = 10)
p + ggsave("COVID vs CTRL MF enrich 0.05.pdf",
           width = 8,
           height = 6)

goBP <-
  enrichGO(
    gene = covidList,
    OrgDb = hm,
    keyType = "UNIPROT",
    ont = "BP",
    pvalueCutoff = 1
  )
goBPdf <- as.data.frame(goBP)
write.csv(goBPdf, "COVID vs CTRL BP enrich.csv", row.names = F)
p <- dotplot(filter(goBP,qvalue<0.2), font.size = 10)
p + ggsave("COVID vs CTRL BP enrich 0.2.pdf",
           width = 8,
           height = 6)

goCC <-
  enrichGO(
    gene = covidList,
    OrgDb = hm,
    keyType = "UNIPROT",
    ont = "CC",
    pvalueCutoff = 1
  )
goCCdf <- as.data.frame(goCC)
write.csv(goCCdf, "COVID vs CTRL CC enrich.csv", row.names = F)
p <- dotplot(filter(goCC,qvalue<0.05), font.size = 10)
p + ggsave("COVID vs CTRL CC enrich 0.05.pdf",
           width = 8,
           height = 6)

kegg <-
  enrichKEGG(
    gene = covidList,
    organism = "hsa",
    keyType = "uniprot",
    pvalueCutoff = 1
  )
keggDf <- as.data.frame(kegg)
write.csv(keggDf, "COVID vs CTRL KEGG enrich.csv", row.names = F)
p <- dotplot(filter(kegg,qvalue<0.1), font.size = 10)
p + ggsave("COVID vs CTRL KEGG enrich 0.1.pdf",
           width = 8,
           height = 6)

#### S ####

S <-
  read.csv("Protein with different abundance S vs CTRL.csv")
SList <- unlist(str_split(S$X, ";"))
SList <- unlist(str_split(SList, "\\|", simplify = T)[, 2])

goMF <-
  enrichGO(
    gene = SList,
    OrgDb = hm,
    keyType = "UNIPROT",
    ont = "MF",
    pvalueCutoff = 1
  )
goMFdf <- as.data.frame(goMF)
write.csv(goMFdf, "S vs CTRL MF enrich.csv", row.names = F)
p <- dotplot(slice(goMF,n=1:10), font.size = 10)
p + ggsave("S vs CTRL MF enrich 1-10.pdf",
           width = 8,
           height = 6)

goBP <-
  enrichGO(
    gene = SList,
    OrgDb = hm,
    keyType = "UNIPROT",
    ont = "BP",
    pvalueCutoff = 1
  )
goBPdf <- as.data.frame(goBP)
write.csv(goBPdf, "S vs CTRL BP enrich.csv", row.names = F)
p <- dotplot(slice(goBP,n=1:10), font.size = 10)
p + ggsave("S vs CTRL BP enrich 1-10.pdf",
           width = 8,
           height = 6)

goCC <-
  enrichGO(
    gene = SList,
    OrgDb = hm,
    keyType = "UNIPROT",
    ont = "CC",
    pvalueCutoff = 1
  )
goCCdf <- as.data.frame(goCC)
write.csv(goCCdf, "S vs CTRL CC enrich.csv", row.names = F)
p <- dotplot(slice(goCC,n=1:10), font.size = 10)
p + ggsave("S vs CTRL CC enrich 1-10.pdf",
           width = 8,
           height = 6)

kegg <-
  enrichKEGG(
    gene = SList,
    organism = "hsa",
    keyType = "uniprot",
    pvalueCutoff = 1
  )
keggDf <- as.data.frame(kegg)
write.csv(keggDf, "S vs CTRL KEGG enrich.csv", row.names = F)
p <- dotplot(slice(kegg,n=1:10), font.size = 10)
p + ggsave("S vs CTRL KEGG enrich 1-10.pdf",
           width = 8,
           height = 6)

#### COVID vs S ####

covid <-
  read.csv("Protein with different abundance COVID vs S.csv")
covidList <- unlist(str_split(covid$X, ";"))
covidList <- unlist(str_split(covidList, "\\|", simplify = T)[, 2])

goMF <-
  enrichGO(
    gene = covidList,
    OrgDb = hm,
    keyType = "UNIPROT",
    ont = "MF",
    pvalueCutoff = 1
  )
goMFdf <- as.data.frame(goMF)
write.csv(goMFdf, "COVID vs S MF enrich.csv", row.names = F)
p <- dotplot(slice(goMF,n=1:10), font.size = 10)
p + ggsave("COVID vs S MF enrich 1-10.pdf",
           width = 8,
           height = 6)

goBP <-
  enrichGO(
    gene = covidList,
    OrgDb = hm,
    keyType = "UNIPROT",
    ont = "BP",
    pvalueCutoff = 1
  )
goBPdf <- as.data.frame(goBP)
write.csv(goBPdf, "COVID vs S BP enrich.csv", row.names = F)
p <- dotplot(slice(goBP,n=1:10), font.size = 10)
p + ggsave("COVID vs S BP enrich 1-10.pdf",
           width = 8,
           height = 6)

goCC <-
  enrichGO(
    gene = covidList,
    OrgDb = hm,
    keyType = "UNIPROT",
    ont = "CC",
    pvalueCutoff = 1
  )
goCCdf <- as.data.frame(goCC)
# empty result
# write.csv(goCCdf, "COVID vs S CC enrich.csv", row.names = F)
# p <- dotplot(slice(goCC,n=1:10), font.size = 10)
# p + ggsave("COVID vs S CC enrich 1-10.pdf",
#            width = 8,
#            height = 6)

kegg <-
  enrichKEGG(
    gene = covidList,
    organism = "hsa",
    keyType = "uniprot",
    pvalueCutoff = 1
  )
keggDf <- as.data.frame(kegg)
# empty result
# write.csv(keggDf, "COVID vs S KEGG enrich.csv", row.names = F)
# p <- dotplot(slice(kegg,n=1:10), font.size = 10)
# p + ggsave("COVID vs S KEGG enrich 1-10.pdf",
#            width = 8,
#            height = 6)
```

# sipros result

```{r}
library(ggvenn)
siDf <- fread("Pan_050721.pro.txt", skip = 58)
si <- siDf$ProteinID
si <- si[!str_detect(si, "Rev2_")]
si <- str_replace(si, "\\{", "")
si <- str_replace(si, "\\}", "")
si <- unlist(str_split(si, ","))
mq <- read.csv("proteinGroupsFiltered.csv")
mq <- mq$Majority.protein.IDs
mq <- unlist(str_split(mq, ";"))
# 3953
sum(si %in% mq)
p <- venn.diagram(
  x = list(Sipros = si,
           Maxquant = mq),
  filename = NULL,
  output = F ,
  height = 1200 ,
  width = 1200 ,
  resolution = 30
)
pdf(file = "Sipros Maxquant venn.pdf")
grid.draw(p)
dev.off()

a <- list(Sipros = si,
          Maxquant = mq)
p <- ggvenn(a, c("Sipros", "Maxquant"))
p + ggsave("Sipros Maxquant venn2.pdf")
```

