---
title: "HW1"
author: "Yi Xiong"
date: "2021/1/31"
output: 
  pdf_document:
    toc: true
---

# 1)Generate $\epsilon$ ~t~\~N(0,1) ; IID samples, 1 $\leq$ t $\leq$ 500

## a)

### $\epsilon$~t~(vs)t

```{r}
library(ggplot2)
x<-1:500
set.seed(6666)
y<-rnorm(500)
df<-data.frame(x=x,y=y)
p<-ggplot(df,aes(x=x,y=y))+
  geom_line()+
  xlab("t")+
  ylab(expression(epsilon[t]))
p
```

$\epsilon$~t~ fluctuates around 0

### $\epsilon$~t~ (vs) $\epsilon$~t-1~

```{r}
set.seed(6666)
y<-rnorm(501)
t<-501
df<-data.frame(x=y[1:c(t-1)],y=y[2:t])
p<-ggplot(df,aes(x=x,y=y))+
  geom_point()+
  xlab(expression(epsilon[t-1]))+
  ylab(expression(epsilon[t]))
p
```

it seems no correlation between $\epsilon$~t~ and $\epsilon$~t-1~

### $\epsilon$~t~ (vs) $\epsilon$~t-2~

```{r}
set.seed(6666)
y<-rnorm(502)
t<-502
df<-data.frame(x=y[1:(t-2)],y=y[3:t])
p<-ggplot(df,aes(x=x,y=y))+
  geom_point()+
  xlab(expression(epsilon[t-2]))+
  ylab(expression(epsilon[t]))
p
```

it seems no correlation between $\epsilon$~t~ and $\epsilon$~t-2~

## b)

### mean

```{r}
set.seed(6666)
y<-rnorm(500)
mean(y)
```

### variance

```{r}
var(y)
```

### covariance $\gamma$~k~

```{r}
set.seed(6666)
y<-rnorm(500)
getCov<-function(k){
  x1<-y[1:100]
  x2<-y[(k+1):(100+k)]
  cov(x1,x2)
}
k<-0:100
cov1<-sapply(k,getCov)
dfCov<-data.frame(k=k,cov=cov1)
head(dfCov)
```

### $\rho$~k~

```{r}
p<-cov1/cov1[1]
dfP<-data.frame(k=k,p=p)
head(dfP)
```

### $\gamma$~k~ (vs) k

```{r}
p<-ggplot(dfCov,aes(x=k,y=cov))+
  geom_point()+
  xlab(expression(k))+
  ylab(expression(gamma[k]))
p
```

### $\rho$~k~ (vs) k

```{r}
p<-ggplot(dfP,aes(x=k,y=p))+
  geom_point()+
  xlab(expression(k))+
  ylab(expression(rho[k]))
p
```

## c)

### min

```{r}
set.seed(6666)
y<-rnorm(500)
min(y)
```

### max

```{r}
max(y)
```

### histogram

```{r}
qplot(y,  binwidth=0.1, geom = "histogram", xlab=expression(epsilon[t]), ylab="Frequency") 
```

# 2) Repeat (Q1) on x~t~ when x~t~ is given by

## a) x~t~=2.0+$\epsilon$~t~, x~0~=0

### x~t~(vs)t

```{r}
t<-1:500
set.seed(6666)
y<-rnorm(500)+2.0
df<-data.frame(x=t,y=y)
p<-ggplot(df,aes(x=x,y=y))+
  geom_line()+
  xlab("t")+
  ylab(expression(x[t]))
p
```

x~t~ fluctuates around 2

### x~t~ (vs) x~t-1~

```{r}
set.seed(6666)
y<-rnorm(501)+2.0
t<-501
df<-data.frame(x=y[1:c(t-1)],y=y[2:t])
p<-ggplot(df,aes(x=x,y=y))+
  geom_point()+
  xlab(expression(x[t-1]))+
  ylab(expression(x[t]))
p
```

it seems no correlation between x~t~ and x~t-1~

### x~t~ (vs) x~t-2~

```{r}
set.seed(6666)
y<-rnorm(502)+2.0
t<-502
df<-data.frame(x=y[1:(t-2)],y=y[3:t])
p<-ggplot(df,aes(x=x,y=y))+
  geom_point()+
  xlab(expression(x[t-2]))+
  ylab(expression(x[t]))
p
```
it seems no correlation between x~t~ and x~t-2~


### mean

```{r}
set.seed(6666)
y<-rnorm(501)+2.0
mean(y)
```

### variance

```{r}
var(y)
```

### covariance $\gamma$~k~

```{r}
set.seed(6666)
y<-rnorm(500)+2.0
getCov<-function(k){
  x1<-y[1:100]
  x2<-y[(k+1):(100+k)]
  cov(x1,x2)
}
k<-0:100
cov1<-sapply(k,getCov)
dfCov<-data.frame(k=k,cov=cov1)
head(dfCov)
```

### $\rho$~k~

```{r}
p<-cov1/cov1[1]
dfP<-data.frame(k=k,p=p)
head(dfP)
```

### $\gamma$~k~ (vs) k

```{r}
p<-ggplot(dfCov,aes(x=k,y=cov))+
  geom_point()+
  xlab(expression(k))+
  ylab(expression(gamma[k]))
p
```

### $\rho$~k~ (vs) k

```{r}
p<-ggplot(dfP,aes(x=k,y=p))+
  geom_point()+
  xlab(expression(k))+
  ylab(expression(rho[k]))
p
```

### min

```{r}
set.seed(6666)
y<-rnorm(500)+2.0
min(y)
```

### max

```{r}
max(y)
```

### histogram

```{r}
qplot(y,  binwidth=0.1, geom = "histogram", xlab=expression(x[t]), ylab="Frequency") 
```

## b) x~t~=x~t-1~+$\epsilon$~t~, x~0~=0

### x~t~(vs)t

```{r}
t<-1:500
set.seed(6666)
y<-rnorm(500)
xt<-function(t){
  if (t==0) return(0)
  else return(xt(t-1)+y[t])
}
y<-sapply(1:500,xt)
df<-data.frame(x=t,y=y)
p<-ggplot(df,aes(x=x,y=y))+
  geom_line()+
  xlab("t")+
  ylab(expression(x[t]))
p
```

x~t~ is unstable

### x~t~ (vs) x~t-1~

```{r}
t<-501
set.seed(6666)
y<-rnorm(501)
xt<-function(t){
  if (t==0) return(0)
  else return(xt(t-1)+y[t])
}
y<-sapply(1:t,xt)
df<-data.frame(x=y[1:(t-1)],y=y[2:t])
p<-ggplot(df,aes(x=x,y=y))+
  geom_point()+
  xlab(expression(x[t-1]))+
  ylab(expression(x[t]))
p
```

it seems strong correlation between x~t~ and x~t-1~

### x~t~ (vs) x~t-2~

```{r}
t<-502
set.seed(6666)
y<-rnorm(502)
xt<-function(t){
  if (t==0) return(0)
  else return(xt(t-1)+y[t])
}
y<-sapply(1:t,xt)
df<-data.frame(x=y[1:(t-2)],y=y[3:t])
p<-ggplot(df,aes(x=x,y=y))+
  geom_point()+
  xlab(expression(x[t-2]))+
  ylab(expression(x[t]))
p
```
it seems strong correlation between x~t~ and x~t-2~


### mean

```{r}
set.seed(6666)
y<-rnorm(500)
xt<-function(t){
  if (t==0) return(0)
  else return(xt(t-1)+y[t])
}
y<-sapply(1:500,xt)
mean(y)
```

### variance

```{r}
var(y)
```

### covariance $\gamma$~k~

```{r}
set.seed(6666)
y<-rnorm(1000)
y<-cumsum(y)
getCov<-function(k){
  x1<-y[1:500]
  x2<-y[(k+1):(500+k)]
  cov(x1,x2)
}
k<-0:100
cov1<-sapply(k,getCov)
dfCov<-data.frame(k=k,cov=cov1)
head(dfCov)
```

### $\rho$~k~

```{r}
p<-cov1/cov1[1]
dfP<-data.frame(k=k,p=p)
head(dfP)
```

### $\gamma$~k~ (vs) k

```{r}
p<-ggplot(dfCov,aes(x=k,y=cov))+
  geom_line()+
  xlab(expression(k))+
  ylab(expression(gamma[k]))
p
```

### $\rho$~k~ (vs) k

```{r}
p<-ggplot(dfP,aes(x=k,y=p))+
  geom_line()+
  xlab(expression(k))+
  ylab(expression(rho[k]))
p
```

### min

```{r}
set.seed(6666)
y<-rnorm(500)
y<-cumsum(y)
min(y)
```

### max

```{r}
max(y)
```

### histogram

```{r}
qplot(y,  binwidth=1, geom = "histogram", xlab=expression(x[t]), ylab="Frequency")
```

# 3) Verify convariance function $\mathit{f(i,j)}$ is non-negtive definite

```{r}
# define xt~N(0,1)
set.seed(6666)
xt<-rnorm(500)
# each column of xt is x(i) or x(j)
xt<-matrix(xt,ncol = 20)
# verify convariance function f(i,j) is non-negtive definite
totalSum<-0
for (i in 1:20){
  for (j in 1:20){
    totalSum<-totalSum+(t(xt[,i])*cov(xt[,i],xt[,j]))%*%matrix(xt[,j])
  }
}
# TRUE
totalSum>0

# define xt~N(0,1)
set.seed(6666)
xt<-rnorm(500)
# define yt=y(t-1)+xt
yt<-cumsum(xt)
# each column of yt is y(i) or y(j)
yt<-matrix(yt,ncol = 20)
# verify convariance function f(i,j) is non-negtive definite
totalSum<-0
for (i in 1:20){
  for (j in 1:20){
    totalSum<-totalSum+(t(yt[,i])*cov(yt[,i],yt[,j]))%*%matrix(yt[,j])
  }
}
# TRUE
totalSum>0
```


