---
title: "HW2_Yi_Xiong"
author: "Yi Xiong"
date: "2021/2/14"
output: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1) Generate one set of 100 samples of x~t~\~iid N(0,1), and compute its mean $\overline{x}$, and variance $\widehat{\sigma }^{2}$

```{r}
set.seed(6666)
x<-rnorm(100)
# mean
(x_mean<-mean(x))
# variance
(x_var<-var(x))
```

## a) Repeat this experiment 100 times and obtain 100 estimates of $\overline{x}$, and $\widehat{\sigma }^{2}$

```{r}
xt <- data.frame(T = 1:100)
for (i in 1:100) {
  xt <- cbind(xt, rnorm(100))
}
colnames(xt)[-1] <- paste0("S", 1:100)
# mean
head(xt_mean <- colMeans(xt[, -1]))
# variance
head(xt_var <- apply(xt[, -1], 2, var))
```

## b) Compute the mean, variance of $\overline{x}_{i}$, for $1\leq i\leq 100$

```{r}
# mean
(xt_mean_mean <- mean(xt_mean))
# variance
(xt_mean_var <- var(xt_mean))
```

## c) Compute the mean, variance of $\widehat{\sigma_{i}}^{2}$, for $1\leq i\leq 100$

```{r}
# mean
(xt_var_mean <- mean(xt_var))
# variance
(xt_var_var <- var(xt_var))
```

# 2) Generate $\varepsilon _{t}$\~iid N(0,1) for $1\leq t\leq 500$

```{r}
t<-500
set.seed(6666)
x<-rnorm(500)
```

## a) Define $x_{t}=\varepsilon _{t}+0.5\varepsilon _{t-1}$ called MA(1) process

```{r}
xt<-c(x[1])
for (t in 2:500){
  xt[t]<-x[t]+0.5*x[t-1]
}
```

## b) Plot x~t~ (vs)x~t-1~ , x~t~ (vs) x~t-2~ ,x~t~ (vs) ~t~

```{r}
library(ggplot2)
# t vs t-1
df<-data.frame(x=xt[1:(t-1)],y=xt[2:t])
p<-ggplot(df,aes(x=x,y=y))+
  geom_point()+
  xlab(expression(x[t-1]))+
  ylab(expression(x[t]))
p
# t vs t-2
df<-data.frame(x=xt[1:(t-2)],y=xt[3:t])
p<-ggplot(df,aes(x=x,y=y))+
  geom_point()+
  xlab(expression(x[t-2]))+
  ylab(expression(x[t]))
p
# xt vs t
df<-data.frame(x=1:t,y=xt)
p<-ggplot(df,aes(x=x,y=y))+
  geom_line()+
  xlab(expression(t))+
  ylab(expression(x[t]))
p
```

## c) Compute the mean, variance and autocorrelation of x~t~ and plot

```{r}
# mean
(xt_mean<-mean(xt))
# var
(xt_var<-var(xt))
# autocorrelation
getCov<-function(k){
  x1<-xt[1:100]
  x2<-xt[(k+1):(100+k)]
  cov(x1,x2)
}
k<-0:100
cov1<-sapply(k,getCov)
dfCov<-data.frame(k=k,cov=cov1)
p<-ggplot(dfCov,aes(x=k,y=cov))+
  geom_point()+
  xlab(expression(k))+
  ylab(expression(gamma[k]))
p

p<-cov1/cov1[1]
dfP<-data.frame(k=k,p=p)
p<-ggplot(dfP,aes(x=k,y=p))+
  geom_point()+
  xlab(expression(k))+
  ylab(expression(rho[k]))
p
```

# 3) Plot $N\left( 0,\sigma ^{2}\right)$ for $\sigma ^{2}$ = 0.1, 0.5, 1.0, 2.0, 5.0 on the same plot and comment

```{r}
# ggplot() + xlim(-5, 5) + geom_function(
#   fun = function(x)
#     dnorm(x, sd = sqrt(0.1)),
#   colour = "red"
# ) + geom_function(
#   fun = function(x)
#     dnorm(x, sd = sqrt(0.5)),
#   colour = "blue"
# ) + geom_function(
#   fun = function(x)
#     dnorm(x, sd = sqrt(1)),
#   colour = "green"
# ) + geom_function(
#   fun = function(x)
#     dnorm(x, sd = sqrt(2)),
#   colour = "yellow"
# ) + geom_function(
#   fun = function(x)
#     dnorm(x, sd = sqrt(5)),
#   colour = "orange"
# ) +
#   ylab("Density")+
#   scale_colour_manual(values=cols)

x1<-seq(-5,5,length.out=1001)
d<-sapply(c(0.1,0.5,1,2,5), function(x)dnorm(x1,sd = sqrt(x)))
d<-data.frame(x=rep(x1,times=5),density=c(d),sigma=rep(c(0.1,0.5,1,2,5),each=1001))
d$sigma<-as.factor(d$sigma)
p<-ggplot(d,aes(x=x,y=density,color=sigma))+
  geom_line()+
  xlab("")+
  ylab("Density")+
  scale_color_discrete(name = expression(sigma^2)) 
p
```

As $\sigma ^{2}$ decreases, the tail gets thiner, the overall spread decreases, and the peak at $\mu$ = 0 increases, 
